AFQ-Insight

### 2019-12-01 reports/manuscript/abstract.tex
More references and edits throughout.

The white matter contains long-range connections between different brain
regions. Tractometry uses diffusion-weighted magnetic resonance imaging (dMRI)
data to quantify tissue properties along the trajectories of these connections
in the human brain in vivo\cite{yeatman2012tract}. In previous work, the results
of tractometry were usually analyzed using mass univariate approaches: group
comparisons or regression models computed separately for each point along every
one of the tracts.  Alternatively, tissue properties such as fractional
anisotropy (FA) and mean diffusivity (MD) were computed for a specific tract of
interest based on an a priori hypothesis. In the present work, we developed a
method based on the sparse group lasso\cite{simon2013sparse} that takes into
account tissue properties measured in all of the tracts, and selects informative
features by enforcing sparsity, both at the level of individual tracts and
tissue properties, but also across the entire set of tracts and all of the
measured tissue properties. The sparsity penalties for each of these constraints
is identified using a nested cross-validation scheme. Using data from a previous
study that measured dMRI in patients with amyotrophic lateral sclerosis (ALS)
and matched controls\cite{sarica2017corticospinal}, we demonstrate that this
method is accurate, exceeding the previous results, that were based on a priori
feature selection, in classifying patients and controls, with ~84\% accuracy, an
area under the receiver operating characteristic curve of 0.93, and an average
precision of 0.95. Moreover, our method automatically identifies as important
for this classification the parts of the white matter known to be affected by
ALS within the corticospinal tract. In a regression setting, data from another
previous study\cite{yeatman2014lifespan} can be used to accurately predict
``brain age''. Thus, this multivariate analysis approach both (a) achieves high
cross-validated accuracy for precision medicine applications of dMRI data and
(b) identifies relevant features of brain anatomy to further our neuroscientific
understanding of clinical disorders.

>>>>

The white matter contains long-range connections between different brain
regions. Tractometry uses diffusion-weighted magnetic resonance imaging (dMRI)
data to quantify tissue properties, such as fractional anisotropy (FA) and mean
diffusivity (MD), along the trajectories of these connections in the human brain
\emph{in vivo}\cite{yeatman2012tract}. Statistical inference from tractometry
data usually either averages these quantities along the length of the tract in
each individual, or performs analysis point-by-point along each tract, with
group comparisons or regression models computed separately for each point along
every one of the tracts. Alternatively, tissue properties are computed for a
specific tract of interest based on an a priori hypothesis. In the present work,
we developed a method based on the sparse group lasso (SGL)
\cite{simon2013sparse} that takes into account tissue properties measured along
the tracts, and selects informative features by enforcing sparsity, both at the
level of individual tracts and tissue properties, but also across the entire set
of tracts and all of the measured tissue properties. The sparsity penalties for
each of these constraints is identified using a nested cross-validation scheme
that guards against over-fitting and simultaneously identifies the correct level
of sparsity, both at the group level and overall. We demonstrate the accuracy of
the method in two settings: In a regression setting, dMRI is used to accurately
predict ``brain age'' \cite{yeatman2014lifespan, Brown2012-so}. In a
classification setting, patients with amyotrophic lateral sclerosis (ALS) are
accurately distinguished from matched controls\cite{sarica2017corticospinal}. In
addition, SGL automatically identifies FA in the corticospinal tract as
important for this classification -- correctly finding the parts of the white
matter known to be affected by ALS. \ariel{Still need a punchy ending sentence
here}

### 2019-12-01 reports/manuscript/sections/results.tex
Removes todo list already addressed.
More references and edits throughout.

To test the performance of SGL in tractometry data in a continuous regression
task, we focus here on the prediction of biological age from the tractometry
data. Prediction of ``brain age'' is a commonly undertaken task. This is both
because it operates on a natural scale, with meaningful and easily understood
units, as well as because predictions of brain age, and deviations from accurate
prediction are diagnostic of overall brain age \ariel{citation needed}. The WH
dataset used here contains data from 76 healthy subjects, ranging between 6
years and 50 years of age \cite{yeatman2014lifespan}. In this case, biological
age was used as the predicted variable ($y$ in equation \ref{eq:lm}). SGL was
fit to tractometry-extracted features: FA and MD in 20 major brain tracts, with
each tract divided into 100 nodes. To evaluate the fit of the model, we used a
nested cross-validation procedure. In this procedure, batches of subjects are
held out. For each batch (or fold), the model is fully fit without this data.
Then, once the parameters are fixed, the model is inverted to predict the ages
of held out subjects based on the linear coeffiecients and the static
non-linearity. data is held out, and a model age was predicted using a model
that was fit without that subject. This scheme automatically finds the right
level of regularization and fits the coefficients to the ill-posed linear model,
while guarding against overfitting. SGL accurately predicts the age of the
subjects in this procedure, with a mean absolute error of 3.6 years (Figure
\ref{fig:regress-results}, left panel). This is lower than the results of a
recent study that predicted age in a large sample, based on diffusion MRI
features \cite{Richard2018-ux}. Nevertheless, older subjects have higher
residual variance, reflecting the automatically-chosen log-transformation and
implying that brain age becomes more difficult to predict as we age
chronologically (\ref{fig:regress-result}, right panel). The model weights are
distributed over many different tracts and dMRI tissue properties (Figure
\ref{fig:regress-beta} left). This demonstrates that SGL is not coerced to
produce overly sparse results when a more accurate model requires a dense
selection of features. Furthermore, looking closer at a selection of tracts
where high coefficients are found demonstrates that diffusion properties (FA, in
this case) are different in different age groups in parts of the tracts where
these higher coefficients are found (Figure \ref{regress-beta} right).

\begin{figure}[!h]
    \centering
    \includegraphics[width=0.8\textwidth]{regression_residuals.png}
    \caption{{\bf Predicting age with tractometry and SGL.} Left: The predicted age of each individual (on the abscissa) and true age (on the ordinate), from the test splits (i.e., when each subject's data was held out in fitting the model); an accurate prediction falls close to the $y=x$ line (dashed). The mean absolute error in this case is 3.6 years and, the coefficient of determination $R^2=0.3$. Right: Standardized residuals (on the abscissa) as a function of the true age (on the ordinate). Predictions are generally more accurate for younger individuals.
    }
    \label{fig:regress-results}
\end{figure}

\begin{figure}[!h]
    \centering
    \includegraphics[width=0.3\textwidth]{regression_beta_annotated.pdf}
    \includegraphics[width=0.67\textwidth]{regression_tract_profiles.pdf}
    \caption{{\bf Feature importance for predicting age from tractometry.} Left: A skeletonized display of the main brain tracts analyzed, with anterior facing up, and right hemisphere on the right. The $\hat{\beta}$ coefficients displayed in blue (negative) to red (positive) are for measurements of FA along the length of the tracts. The left cingulum cingulate (A) and forceps minor (B) are highlighted. Right: the FA (in shades of blue and green) and the $beta$ coefficients (dashed) in (A) left cingulum and (B) forceps minor.
    }
    \label{fig:regress-beta}
\end{figure}

\begin{itemize}
  \item Regression
    \begin{itemize}
      \item Lifespan maturation age regression
      \item Insert figure showing sparsity pattern
      \item Insert figure showing weights as they relate to tract differences visible in the browser
    \end{itemize}
\end{itemize}
\begin{itemize}
  \item Failures
    \begin{itemize}
      \item Hopefully, the failures are common to both regression
        and classification so we can include them here in there own
        subsection.
      \item Insert figure demonstrating failure cases
    \end{itemize}
\end{itemize}

>>>>

To test the performance of SGL in tractometry data in a continuous regression
task, we focus here on the prediction of biological age from the tractometry
data. Prediction of ``brain age'' is a commonly undertaken task. This is both
because it operates on a natural scale, with meaningful and easily understood
units, as well as because predictions of brain age, and deviations from accurate
prediction are diagnostic of overall brain health (for a recent review, see
\cite{Cole2019-rz}). The WH dataset used here contains data from 76 healthy
subjects, ranging between 6 years and 50 years of age
\cite{yeatman2014lifespan}. In this case, biological age was used as the
predicted variable ($y$ in equation \ref{eq:lm}). SGL was fit to
tractometry-extracted features: FA and MD in 20 major brain tracts, with each
tract divided into 100 nodes. To evaluate the fit of the model, we used a nested
cross-validation procedure. In this procedure, batches of subjects are held out.
For each batch (or fold), the model is fully fit without this data. Then, once
the parameters are fixed, the model is inverted to predict the ages of held out
subjects based on the linear coeffiecients and the static non-linearity. data is
held out, and a model age was predicted using a model that was fit without that
subject. This scheme automatically finds the right level of regularization and
fits the coefficients to the ill-posed linear model, while guarding against
overfitting. SGL accurately predicts the age of the subjects in this procedure,
with a mean absolute error of 3.6 years (Figure \ref{fig:regress-results}, left
panel). This is lower than the results of a recent study that predicted age in a
large sample, based on diffusion MRI features \cite{Richard2018-ux}.
Nevertheless, older subjects have higher residual variance, reflecting the
automatically-chosen log-transformation and implying that brain age becomes more
difficult to predict as we age chronologically (\ref{fig:regress-results}, right
panel). The model weights are distributed over many different tracts and dMRI
tissue properties (Figure \ref{fig:regress-beta} left). This demonstrates that
SGL is not coerced to produce overly sparse results when a more accurate model
requires a dense selection of features. Furthermore, looking closer at a
selection of tracts where high coefficients are found demonstrates that
diffusion properties (FA, in this case) are different in different age groups in
parts of the tracts where these higher coefficients are found (Figure
\ref{fig:regress-beta} right).

\begin{figure}[!h]
    \centering
    \includegraphics[width=0.8\textwidth]{regression_residuals.png}
    \caption{{\bf Predicting age with tractometry and SGL.} Left: The predicted
    age of each individual (on the abscissa) and true age (on the ordinate),
    from the test splits (i.e., when each subject's data was held out in fitting
    the model); an accurate prediction falls close to the $y=x$ line (dashed).
    The mean absolute error in this case is 3.6 years and, the coefficient of
    determination $R^2=0.3$. Right: Standardized residuals (on the abscissa) as
    a function of the true age (on the ordinate). Predictions are generally more
    accurate for younger individuals.}
    \label{fig:regress-results}
\end{figure}

\begin{figure}[!h]
    \centering
    \includegraphics[width=0.3\textwidth]{regression_beta_annotated.pdf}
    \includegraphics[width=0.67\textwidth]{regression_tract_profiles.pdf}
    \caption{{\bf Feature importance for predicting age from tractometry.} Left:
    A skeletonized display of the main brain tracts analyzed, with anterior
    facing up, and right hemisphere on the right. The $\hat{\beta}$ coefficients
    displayed in blue (negative) to red (positive) are for measurements of FA
    along the length of the tracts. The left cingulum cingulate (A) and forceps
    minor (B) are highlighted. Right: the FA (in shades of blue and green) and
    the $beta$ coefficients (dashed) in (A) left cingulum and (B) forceps
    minor.}
    \label{fig:regress-beta}
\end{figure}
